---
title: "Preliminary Analysis"
author: "Andie Creel"
date: "6/15/2021"
output: 
  pdf_document:
    number_sections: true
    extra_dependencies: ["float"]

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'Exploratory_Output/regressions/preliminary_analysis.pdf')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")


library(vroom)
library(dplyr)
library(tidyr)
library(stringr)
library(usmap)
library(tidycensus)
library(ggplot2)
library(fixest)
library(kableExtra)
library(gtools)
library(forcats)
library("wesanderson")
options(scipen=99999)
library(janitor)

myWorking <- vroom("Datasets/clean_data/lwcf_tws_projects_onlyACS.csv")

```


```{r,  echo=FALSE}
# additional data cleaning ------------

# Removing counties whose FIPSs code I didn't identify 
myWorking <- myWorking %>%
  filter(!is.na(county_fips)) %>%
  distinct()

#removing grants we don't know the size of or that are zero 
myWorking <- myWorking %>%
  filter(!is.na(amount)) %>%
  filter(amount != 0)

myPercentiles <- quantile(myWorking$amount, probs = c(0, 0.05, .1, .25, .5, .75, .9, .95, 1))

#dropping parks in the lowest 5% and in the highest 5% (about the 95th percentile)

myWorking <- myWorking %>%
  filter(amount > myPercentiles["5%"]) %>%
  filter(amount < myPercentiles["95%"])


# Adding metro, nonmetro and rural variable
myUrbanKey <- myWorking %>%
  dplyr::select(RUCC_2013, Description) %>%
  distinct()

myWorking <- myWorking %>%
  mutate(metro = if_else(RUCC_2013 == 1 | RUCC_2013 == 2 | RUCC_2013 == 3, 1, 0)) %>%
  mutate(nonmetro = if_else(RUCC_2013 == 4 | RUCC_2013 == 5 | RUCC_2013 == 6, 1, 0)) %>%
  mutate(rural = if_else(RUCC_2013 == 7 | RUCC_2013 == 8 | RUCC_2013 == 9, 1, 0))


# getting ready for deciles
myWorking <- myWorking %>%
  mutate(type = as.factor(type)) %>%
  mutate(black_dec = quantcut(black_pct, q = 10, na.rm=TRUE)) %>%
  mutate(white_dec = quantcut(white_pct, q = 10, na.rm=TRUE)) %>%
  mutate(asian_dec = quantcut(asian_pct, q = 10, na.rm=TRUE)) %>%
  mutate(native_dec = quantcut(native_pct, q = 10, na.rm=TRUE)) %>%
  mutate(hispanic_dec = quantcut(hispanic_pct, q = 10, na.rm=TRUE)) %>%
  mutate(poverty_dec = quantcut(inc_below_pov_pct, q = 10, na.rm=TRUE)) %>%
  mutate(black_dec = fct_recode(black_dec, 
           ten = levels(black_dec)[10])) %>%
  mutate(white_dec = fct_recode(white_dec, 
           ten = levels(white_dec)[10])) %>%  
  mutate(asian_dec = fct_recode(asian_dec, 
           ten = levels(asian_dec)[10])) %>%  
 mutate(native_dec = fct_recode(native_dec, 
           ten = levels(native_dec)[10])) %>%  
  mutate(hispanic_dec = fct_recode(hispanic_dec, 
           ten = levels(hispanic_dec)[10])) %>%
  mutate(poverty_dec = fct_recode(poverty_dec, 
           ten = levels(poverty_dec)[10]))

# making 0/1 for counties in the highest decile for different races and ethnicities 
myWorking <- myWorking %>%
  mutate(white_county = if_else(white_dec == "ten", 1, 0)) %>%
  mutate(black_county = if_else(black_dec == "ten", 1, 0)) %>%
  mutate(asian_county = if_else(asian_dec == "ten", 1, 0)) %>%
  mutate(native_county = if_else(native_dec == "ten", 1, 0)) %>%
  mutate(hispanic_county = if_else(hispanic_dec == "ten", 1, 0)) 


# Removing useless variables 
myWorking <- myWorking %>%
  dplyr::select(-c(objectid, county, grantidelement, grantelementtitle, grantsponsor, congdistrict, relate))

# other outcome variable -------------------------------------
# annual quantity of grants

 myCounts <- myWorking %>% 
  group_by(fiscal_year, county_fips) %>%
  tally() %>%
  dplyr::rename(annual_quantity = n)
 
 myWorking_quantity <- left_join(myWorking, myCounts , by = c("fiscal_year", "county_fips")) %>%
   dplyr::select(-type, -amount) %>%
   distinct()

# annual amount of funding
mySum <- myWorking %>%
   group_by(fiscal_year, county_fips) %>%
    dplyr::summarise(annual_amount = sum(amount))

  
myWorking_quantity <- left_join(myWorking_quantity, mySum , by = c("fiscal_year", "county_fips"))


```


```{r,  echo=FALSE, message=FALSE}
myWorking %>%
  tabyl(type) %>%
  kbl() %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

```



# Regressing each Race/Ethnicity Seperately 

## Perctage Regressions 

### Amount of Funding per Reward

```{r,  echo=FALSE, message=FALSE}
# White
one <- feols(data = myWorking, log(amount) ~ white_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_pct + med_income_house + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "White Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Black

one <- feols(data = myWorking, log(amount) ~ black_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ black_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ black_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ black_pct + med_income_house + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Black Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))  

# asian

one <- feols(data = myWorking, log(amount) ~ asian_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ asian_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ asian_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ asian_pct + med_income_house + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Asian Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# native

one <- feols(data = myWorking, log(amount) ~ native_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ native_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ native_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ native_pct + med_income_house + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Native Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# hispanic

one <- feols(data = myWorking, log(amount) ~ hispanic_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ hispanic_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ hispanic_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ hispanic_pct + med_income_house + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Hispanic Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# poverty

one <- feols(data = myWorking, log(amount) ~ inc_below_pov_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ inc_below_pov_pct + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ inc_below_pov_pct  + noCollegeDegree_pct + metro + nonmetro + rural + type | state_fips + fiscal_year)

etable(one,
       two,
       three,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Income Below Poverty Level Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


```


<!-- * poverty and no college degree are highly correlated  -->


### Annual Quantity of Grants

```{r,  echo=FALSE, message=FALSE}


# White
one <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + med_income_house  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year )
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "White Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Black

one <- feols(data = myWorking_quantity, log(annual_quantity) ~ black_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ black_pct + med_income_house  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ black_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year )
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ black_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Black Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# asian

one <- feols(data = myWorking_quantity, log(annual_quantity) ~ asian_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ asian_pct + med_income_house  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ asian_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year )
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ asian_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Asian Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# native

one <- feols(data = myWorking_quantity, log(annual_quantity) ~ native_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ native_pct + med_income_house  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ native_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year )
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ native_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Native Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# hispanic

one <- feols(data = myWorking_quantity, log(annual_quantity) ~ hispanic_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ hispanic_pct + med_income_house  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ hispanic_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year )
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ hispanic_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Hispanic Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))  


# poverty

one <- feols(data = myWorking_quantity, log(annual_quantity) ~ inc_below_pov_pct  + population | state_fips + fiscal_year )
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips + fiscal_year )
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ inc_below_pov_pct  + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Under Poverty Level Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


```


### Annual Amount of Grants Funding 

```{r,  echo=FALSE, message=FALSE}


# White
one <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + med_income_house  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "White Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Black

one <- feols(data = myWorking_quantity, log(annual_amount) ~ black_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ black_pct + med_income_house  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ black_pct + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myWorking_quantity, log(annual_amount) ~ black_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Black Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# asian

one <- feols(data = myWorking_quantity, log(annual_amount) ~ asian_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ asian_pct + med_income_house  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ asian_pct + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myWorking_quantity, log(annual_amount) ~ asian_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Asian Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position"))  

# native

one <- feols(data = myWorking_quantity, log(annual_amount) ~ native_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ native_pct + med_income_house  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ native_pct + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myWorking_quantity, log(annual_amount) ~ native_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Native Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# hispanic

one <- feols(data = myWorking_quantity, log(annual_amount) ~ hispanic_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ hispanic_pct + med_income_house  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ hispanic_pct + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myWorking_quantity, log(annual_amount) ~ hispanic_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Hispanic Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# poverty

one <- feols(data = myWorking_quantity, log(annual_amount) ~ inc_below_pov_pct  + population | state_fips )
two <- feols(data = myWorking_quantity, log(annual_amount) ~ inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips )
three <- feols(data = myWorking_quantity, log(annual_amount) ~ inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Under Poverty Level Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


```


## Decile Regressions
```{r,  echo=FALSE, message=FALSE}
# White
one <- feols(data = myWorking, log(amount) ~ white_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_county + med_income_house + noCollegeDegree_pct+ type  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "White Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Black

one <- feols(data = myWorking, log(amount) ~ black_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ black_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ black_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ black_county + med_income_house + noCollegeDegree_pct+ type  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Black Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# asian

one <- feols(data = myWorking, log(amount) ~ asian_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ asian_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ asian_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ asian_county + med_income_house + noCollegeDegree_pct+ type  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Asian Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# native

one <- feols(data = myWorking, log(amount) ~ native_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ native_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ native_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ native_county + med_income_house + noCollegeDegree_pct+ type  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Native Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 
# hispanic

one <- feols(data = myWorking, log(amount) ~ hispanic_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ hispanic_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ hispanic_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ hispanic_county + med_income_house + noCollegeDegree_pct+ type  + metro + nonmetro + rural| state_fips + fiscal_year )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Hispanic Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


```


Trend I'm starting to see: college degree and income are a strong indicator of a county getting less (on most metrics I've checked so far)

# Regressing together 

## Percentage 

### Amount per Individual Award

```{r,  echo=FALSE, message=FALSE}
#regressing all races together

# Median income 
one <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))






# poverty percentage 
one <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

* income below poverty level and no college degree is highly correlated 

### Annual Amount
```{r,  echo=FALSE, message=FALSE}
#regressing all races together


# Median income 
one <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

### Annual Quantity 

```{r,  echo=FALSE, message=FALSE}
#regressing all races together


# Median income 
one <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Percentage Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

When aggregated annual, non-rural areas are receiving more total funds and a larger quantity of funds

** using log(annual quanity) as an outcome variable doesn't seem very valid. **

## Decile 

### Amount per Individual Award

```{r,  echo=FALSE, message=FALSE}
#regressing all races together

# Median income 
one <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type + population | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + type + population| state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type + population| state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


```



### Annual Amount
```{r,  echo=FALSE, message=FALSE}
#regressing all races together


# Median income 
one <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

### Annual Quantity 

```{r,  echo=FALSE, message=FALSE}
#regressing all races together


# Median income 
one <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips + fiscal_year)
two <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct  + population| state_fips + fiscal_year)
three <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips + fiscal_year)
four <- feols(data = myWorking_quantity, log(annual_quantity) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

** annual quantity does not result in normally distributed errors**


# Per Capita: aggregated over years


```{r,  echo=FALSE, message=FALSE}

myAgg <- vroom("Datasets/clean_data/aggregate_lwcf_tws_projects_onlyACS.csv")

# additional data cleaning ------------

# Removing counties whose FIPSs code I didn't identify 
myAgg <- myAgg %>%
  dplyr::filter(!is.na(county_fips)) %>%
  dplyr::rename(total_amount = `Total LWCF Dollars`) %>%
  dplyr::rename(total_quantity = `Total Projects`) %>%
  dplyr::rename(total_amount_per_cap = `Per Capita LWCF Spending`) %>%
  dplyr::mutate(total_quantity_per_cap = total_quantity/`Population Estimate`)


myAgg <- myAgg %>%
  mutate(metro = if_else(RUCC_2013 == 1 | RUCC_2013 == 2 | RUCC_2013 == 3, 1, 0)) %>%
  mutate(nonmetro = if_else(RUCC_2013 == 4 | RUCC_2013 == 5 | RUCC_2013 == 6, 1, 0)) %>%
  mutate(rural = if_else(RUCC_2013 == 7 | RUCC_2013 == 8 | RUCC_2013 == 9, 1, 0))


# getting ready for deciles
myAgg <- myAgg %>%
  mutate(black_dec = quantcut(black_pct, q = 10, na.rm=TRUE)) %>%
  mutate(white_dec = quantcut(white_pct, q = 10, na.rm=TRUE)) %>%
  mutate(asian_dec = quantcut(asian_pct, q = 10, na.rm=TRUE)) %>%
  mutate(native_dec = quantcut(native_pct, q = 10, na.rm=TRUE)) %>%
  mutate(hispanic_dec = quantcut(hispanic_pct, q = 10, na.rm=TRUE)) %>%
  mutate(poverty_dec = quantcut(inc_below_pov_pct, q = 10, na.rm=TRUE)) %>%
  mutate(black_dec = fct_recode(black_dec, 
           ten = levels(black_dec)[10])) %>%
  mutate(white_dec = fct_recode(white_dec, 
           ten = levels(white_dec)[10])) %>%  
  mutate(asian_dec = fct_recode(asian_dec, 
           ten = levels(asian_dec)[10])) %>%  
 mutate(native_dec = fct_recode(native_dec, 
           ten = levels(native_dec)[10])) %>%  
  mutate(hispanic_dec = fct_recode(hispanic_dec, 
           ten = levels(hispanic_dec)[10])) %>%
  mutate(poverty_dec = fct_recode(poverty_dec, 
           ten = levels(poverty_dec)[10]))

# making 0/1 for counties in the highest decile for different races and ethnicities 
myAgg <- myAgg %>%
  mutate(white_county = if_else(white_dec == "ten", 1, 0)) %>%
  mutate(black_county = if_else(black_dec == "ten", 1, 0)) %>%
  mutate(asian_county = if_else(asian_dec == "ten", 1, 0)) %>%
  mutate(native_county = if_else(native_dec == "ten", 1, 0)) %>%
  mutate(hispanic_county = if_else(hispanic_dec == "ten", 1, 0)) 

```

## Decile Regressions 

### Aggregate total amount per capita

```{r,  echo=FALSE, message=FALSE}
#regressing all races together

# Median income 
one <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county   | state_fips )
two <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house  | state_fips )
three <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  | state_fips )
four <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county   | state_fips )
two <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct  | state_fips )
three <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  | state_fips )
four <- feols(data = myAgg, log(total_amount_per_cap) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips )


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

### Aggregate total amount
```{r,  echo=FALSE, message=FALSE}
#regressing all races together

# Median income 
one <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips )
two <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house  + population| state_fips )
three <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# poverty percentage 
one <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county  + population | state_fips )
two <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct  + population| state_fips )
three <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + population| state_fips )
four <- feols(data = myAgg, log(total_amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct  + metro + nonmetro + rural| state_fips )


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

my_R <- myWorking %>%
  dplyr::filter(type == "R ")



```


We get sign flipping when we look at his per capita vs when we control for populaiton as an independent variable. This is done with all years aggregated though. TO-DO: run this dis-aggregated by year


# Per Capita: Not aggregated over time (yay)

## Decile

```{r,  echo=FALSE, message=FALSE}

myWorking_perCap <- myWorking %>%
  dplyr::mutate(amount = amount / population)


#regressing all races together

# Median income 
one <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type  | state_fips + fiscal_year)
two <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + type | state_fips + fiscal_year)
three <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type  | state_fips + fiscal_year)
two <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + type | state_fips + fiscal_year)
three <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

### no year FE

```{r,  echo=FALSE, message=FALSE}

myWorking_perCap <- myWorking %>%
  dplyr::mutate(amount = amount / population)


#regressing all races together

# Median income 
one <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type  | state_fips )
two <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + type | state_fips )
three <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type | state_fips )
four <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + type  | state_fips )
two <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + type | state_fips )
three <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type | state_fips )
four <- feols(data = myWorking_perCap, log(amount) ~ white_county + black_county + asian_county + hispanic_county +native_county + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips )


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

## Percentage

```{r,  echo=FALSE, message=FALSE}


myWorking_perCap <- myWorking %>%
  dplyr::mutate(amount = amount / population)

#regressing all races together

# Median income 
one <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips + fiscal_year)
two <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + type | state_fips + fiscal_year)
three <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips + fiscal_year)
two <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + type | state_fips + fiscal_year)
three <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

### no year FE

```{r,  echo=FALSE, message=FALSE}

myWorking_perCap <- myWorking %>%
  dplyr::mutate(amount = amount / population)


#regressing all races together

# Median income 
one <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips )
two <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + type | state_fips )
three <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type | state_fips )
four <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips )

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips )
two <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + type | state_fips )
three <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type | state_fips )
four <- feols(data = myWorking_perCap, log(amount) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips )


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```

# Doing this as % of the whole 

```{r,  echo=FALSE, message=FALSE}
# getting the whole

myWhole <- myWorking %>%
  dplyr::group_by(fiscal_year) %>%
  dplyr::summarise(year_total_amount = sum(amount))

myWorking <- left_join(myWorking, myWhole, by = "fiscal_year")

myWorking <- myWorking %>%
  dplyr::mutate(amount_percent = amount / year_total_amount)

```


```{r,  echo=FALSE, message=FALSE}

#regressing all races together


# Median income 
one <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + type | state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + med_income_house + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)

etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Median Income Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


# poverty percentage 
one <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + type  | state_fips + fiscal_year)
two <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + type | state_fips + fiscal_year)
three <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type | state_fips + fiscal_year)
four <- feols(data = myWorking, log(amount_percent) ~ white_pct + black_pct + asian_pct + hispanic_pct +native_pct + inc_below_pov_pct + noCollegeDegree_pct + type + metro + nonmetro + rural| state_fips + fiscal_year)


etable(one,
       two,
       three,
       four,
       signifCode = c("***"=0.01, "**"=0.05, "*"=0.10)) %>%
  kbl(booktabs = T, caption = "Poverty Decile Regs") %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

qqnorm(residuals(one))
qqnorm(residuals(two))
qqnorm(residuals(three))
qqnorm(residuals(four))


```











