---
title: "Rural Descriptive Maps"
author: "Andie Creel"
date: "7/9/2021"
output: 
  pdf_document:
    number_sections: true
    extra_dependencies: ["float"]

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'Exploratory_Output/descriptive/rural.pdf')) })
---

Brown colors tend to indicated positive numbers and purple tend to indicate negative numbers. 

```{r, include=FALSE}
#R Markdown stuff
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
options(tinytex.verbose = TRUE)
```

```{r setup, include=FALSE}
library(vroom)
library(tidyr)
library(stringr)
library(usmap)
library(tidycensus)
library(ggplot2)
library(fixest)
library(kableExtra)
library(gtools)
library(forcats)
library("wesanderson")
options(scipen=99999)
library(janitor)
library(dplyr)
# ---- #
library(choroplethr)
library(choroplethrMaps)
library(usdata)
library(maps)
data(fips_codes)
data(continental_us_states)
# --- #



myWorking <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv")

# additional data cleaning ------------

# Removing counties whose FIPSs code I didn't identify 
myWorking <- myWorking %>%
  filter(!is.na(fips)) %>%
  distinct() %>%
  mutate(type = as.factor(type)) %>%
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>%
  filter(!is.na(white_pct)) %>%
  dplyr::mutate(POC_pct = 100 - white_pct)

# Getting quartiles (Sims 2020)

#Getting the quartile for each state each year! 
myQuants <- myWorking %>%
  dplyr::group_by(state_fips, year) %>%
  dplyr::summarise(quarts = list(quantile(rural_pct, probs = c(.25, .5, .75)))) %>%
  unnest_wider(quarts) %>%
  dplyr::rename(rural_25 = `25%`) %>%
  dplyr::rename(rural_50 = `50%`) %>%
  dplyr::rename(rural_75 = `75%`) 

#joining back in
myWorking <- left_join(myWorking, myQuants, by = c("state_fips", "year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking <- myWorking %>%
  mutate(rural_quarts = if_else(rural_pct >= rural_75, "4", "")) %>% # fourth quartile 
  mutate(rural_quarts = if_else(rural_pct < rural_75, "3", rural_quarts)) %>% # third quartile
  mutate(rural_quarts = if_else(rural_pct < rural_50, "2", rural_quarts)) %>% # second quartile
  mutate(rural_quarts = if_else(rural_pct < rural_25, "1", rural_quarts)) %>% # first quartile 
  select(-rural_75, -rural_50, - rural_25) %>%
  mutate(rural_quarts = as.factor(rural_quarts))

```


```{r DATA-MANIPULATION, echo=FALSE, message=FALSE}

# --- data manipulation to get quantity of grants per quartile--- #

# total grants per state 
myTotal <- myWorking %>%
  group_by(state_fips) %>%
  mutate(state_total = sum(got_grant)) %>%
  select(state_fips, state_total) %>%
  distinct()

# number of grants given to first quart counties ever per state
myOne <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(state_fips) %>%
  mutate(one_total = sum(got_grant)) %>%
  select(state_fips, one_total) %>%
  distinct()

# number of grants given to second quart counties ever per state
myTwo <- myWorking %>%
  filter(rural_quarts == "2") %>%
  group_by(state_fips) %>%
  mutate(two_total = sum(got_grant)) %>%
  select(state_fips, two_total) %>%
  distinct()

# number of grants given to third quart counties ever per state
myThree <- myWorking %>%
  filter(rural_quarts == "3") %>%
  group_by(state_fips) %>%
  mutate(three_total = sum(got_grant)) %>%
  select(state_fips, three_total) %>%
  distinct()

# number of grants given to third quart counties ever per state
myFour <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(state_fips) %>%
  mutate(four_total = sum(got_grant)) %>%
  select(state_fips, four_total) %>%
  distinct()

#merge
myRural_mapData <- left_join(myTotal, myOne , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myTwo , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myThree , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myFour , by = "state_fips")
rm(myTotal, myOne, myTwo, myThree, myFour)

#percents 
myRural_mapData <- myRural_mapData %>%
  mutate(one_pct = round(one_total/state_total * 100, 0 )) %>%
  mutate(two_pct = round(two_total/state_total * 100, 0 )) %>%
  mutate(three_pct = round(three_total/state_total * 100 , 0)) %>%
  mutate(four_pct = round(four_total/state_total * 100, 0 )) 
```


**Rural Quartiles**

I calculate the percentage of a county's population that is rural using data from the decennial census. 

I then categorize counties into four quartiles based on the % of the population that is rural. The first quartile categorizes the most urban counties and the fourth quartile categorizes the most rural counties. 

When calculating quartiles, I group by state and decade. I group by state because the LWCF is administered at the state level, therefore managers are making **intrastate** comparisons when deciding how to allocate funds. I group by decade because the % of the population that is rural may change through time. Therefore, a county that had a significant number of rural people in the 1970s may have experienced changes by the 2010s. I group at the decade level because my demographic characteristics come from the decennial census which is collected at the decade level.  


# Percent of Grants Given to Each Quartile 

For each of the four quartiles, I calculate the percentage received of total grants given from 1966 until 2018. This metric is a helpful moment for the total quantity of grants given to the different quartiles, but ignores time trends as well as the amount of funding allocated. 

## Maps

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myRural_mapData <- left_join(myRural_mapData, fips_codes, by = c("state_fips" = "state_code")) 

myRural_mapData <- myRural_mapData %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region))

# --- maps --- #

# First quartile
myOne_map <- myRural_mapData %>%
  select(region, one_pct) %>%
  rename(value = one_pct) %>%
  filter(!is.na(region))

state_choropleth(myOne_map,
  title = "% of grants given to first quartile of rural counties",
  legend = "Percent %",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr")+
  ggtitle(label = "% of total grants given to first quartile rural counties", subtitle =  "(most urban counties)") + 
  labs(fill = "% received by 1st")

# Second quartile
myTwo_map <- myRural_mapData %>%
  select(region, two_pct) %>%
  rename(value = two_pct) %>%
  filter(!is.na(region))

state_choropleth(myTwo_map,
  title = "% of grants given to second quartile of rural counties",
  legend = "Percent %",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "% received by 2nd")


# Third quartile
myThree_map <- myRural_mapData %>%
  select(region, three_pct) %>%
  rename(value = three_pct) %>%
  filter(!is.na(region))

state_choropleth(myThree_map,
  title = "% of grants given to third quartile of rural counties",
  legend = "Percent %",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr")+ 
  labs(fill = "% received by 3rd")

# Four quartile
myFour_map <- myRural_mapData %>%
  select(region, four_pct) %>%
  rename(value = four_pct) %>%
  filter(!is.na(region))

state_choropleth(myFour_map,
  title = "% of grants given to fourth quartile of rural counties",
  legend = "Percent %",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "% received by 4th") + 
    ggtitle(label = "% of grants given to fourth quartile of rural counties", subtitle =  "(counties with most rural)") 

# difference between fourth and first quartile 

myDifference_map <- myRural_mapData %>%
  select(region, one_pct, four_pct) %>%
  mutate(value = four_pct - one_pct) %>%
  filter(!is.na(region))

state_choropleth(myDifference_map,
  title = "Difference in % grants recieved by fourth quartile and first",
  legend = "Difference in % recieved",
  zoom = continental_us_states,
  num_colors = 7) +
  scale_fill_brewer(palette="Purples", direction = -1) + 
  labs(fill = "Difference in % recieved")


```

**Interpretation:** 

* In Texas, the counties in the first quartile for percentage of population that is rural (ie the most urban counties) receive 49-63%% of the quantity of grants given to Montana. 
* In Texas, the counties in the fourth quartile for % of population that is a in rural (counties with the most rural) receive 4-13% of the quantity of grants given to Montana. 
* In Texas, counties in the fourth quartile receive 59-30% less grants than counties in the first quartile (calculated using a simple diference). 

## Table 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#get percent of rural for the state

state_pct_rural <- myWorking %>%
  select(state_fips, rural_pct) %>%
  distinct() %>%
  group_by(state_fips) %>%
  mutate(avg_rural_pct = mean(rural_pct)) %>%
  select(state_fips, avg_rural_pct) %>%
  distinct()

myRural_table <- left_join(myRural_mapData, state_pct_rural, by = "state_fips" )

myRural_table <- myRural_table %>%
  ungroup() %>%
  select(region, one_pct, two_pct, three_pct, four_pct, avg_rural_pct) %>%
  mutate(difference = four_pct - one_pct) %>%
  filter(region != "district of columbia") %>%
  filter(region != "puerto rico") %>%
  arrange(-avg_rural_pct) %>%
  mutate(avg_rural_pct = round(avg_rural_pct, digits = 2)) %>%
  mutate(region = str_to_title(region)) %>%
  select(region, one_pct, two_pct, three_pct, four_pct, difference, avg_rural_pct)


myRural_table %>%
  kable(booktabs = T, caption = "% grants recieved by quartiles ",
      col.names = c("State",
                    "First",
                    "Second", 
                    "Third",
                    "Fourth",
                    "Difference (four minus first)", 
                    "Avg % of county that's in rural")) %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 

# Plot of avg % rural in state and difference between quarts 
ggplot(myRural_table, aes(x=avg_rural_pct, y=difference)) + geom_point()+ geom_smooth(method = "lm") +theme_bw() + 
  xlab("Avg. % of in rural per County") +
  ylab("Difference in fouth and first quartile") + ggtitle("% grants recieved")

```
When looking at the difference in the % of grants received by quartile, there is a slight downward trend in the fourth quartile receiving less than the first as a state's avg % of counties' population in rural increases.

This may indicate that as the % of a counties that is in rural is more observable (because there is more of a rural population to notice) the disparity in the distribution of funds gets more severe. 

## Time Differences: % Recieved 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

myTime_pct <- function(myStart = 1965, myEnd = 2018){
  
  myWorking <- myWorking %>%
    dplyr::filter(year >= myStart) %>%
    dplyr::filter(year < myEnd)

  
  
# Removing counties whose FIPSs code I didn't identify 
myWorking <- myWorking %>%
  filter(!is.na(fips)) %>%
  distinct() %>%
  mutate(type = as.factor(type)) %>%
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>%
  filter(!is.na(white_pct)) %>%
  dplyr::mutate(POC_pct = 100 - white_pct)

# Getting quartiles (Sims 2020)

#Getting the quartile for each state each year! 
myQuants <- myWorking %>%
  dplyr::group_by(state_fips, year) %>%
  dplyr::summarise(quarts = list(quantile(rural_pct, probs = c(.25, .5, .75)))) %>%
  unnest_wider(quarts) %>%
  dplyr::rename(rural_25 = `25%`) %>%
  dplyr::rename(rural_50 = `50%`) %>%
  dplyr::rename(rural_75 = `75%`)

#joining back in
myWorking <- left_join(myWorking, myQuants, by = c("state_fips", "year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking <- myWorking %>%
  mutate(rural_quarts = if_else(rural_pct >= rural_75, "4", "")) %>% # fourth quartile 
  mutate(rural_quarts = if_else(rural_pct < rural_75, "3", rural_quarts)) %>% # third quartile
  mutate(rural_quarts = if_else(rural_pct < rural_50, "2", rural_quarts)) %>% # second quartile
  mutate(rural_quarts = if_else(rural_pct < rural_25, "1", rural_quarts)) %>% # first quartile 
  select(-rural_75, -rural_50, - rural_25) %>%
  mutate(rural_quarts = as.factor(rural_quarts))

# --- data manipulation to get quantity of grants per quartile--- #

# total grants per state 
myTotal <- myWorking %>%
  group_by(state_fips) %>%
  mutate(state_total = sum(got_grant)) %>%
  select(state_fips, state_total) %>%
  distinct()

# number of grants given to first quart counties ever per state
myOne <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(state_fips) %>%
  mutate(one_total = sum(got_grant)) %>%
  select(state_fips, one_total) %>%
  distinct()

# number of grants given to second quart counties ever per state
myTwo <- myWorking %>%
  filter(rural_quarts == "2") %>%
  group_by(state_fips) %>%
  mutate(two_total = sum(got_grant)) %>%
  select(state_fips, two_total) %>%
  distinct()

# number of grants given to third quart counties ever per state
myThree <- myWorking %>%
  filter(rural_quarts == "3") %>%
  group_by(state_fips) %>%
  mutate(three_total = sum(got_grant)) %>%
  select(state_fips, three_total) %>%
  distinct()

# number of grants given to third quart counties ever per state
myFour <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(state_fips) %>%
  mutate(four_total = sum(got_grant)) %>%
  select(state_fips, four_total) %>%
  distinct()

#merge
myRural_mapData <- left_join(myTotal, myOne , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myTwo , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myThree , by = "state_fips")
myRural_mapData <- left_join(myRural_mapData, myFour , by = "state_fips")
rm(myTotal, myOne, myTwo, myThree, myFour)

#percents 
myRural_mapData <- myRural_mapData %>%
  mutate(one_pct = round(one_total/state_total * 100, 0 )) %>%
  mutate(two_pct = round(two_total/state_total * 100, 0 )) %>%
  mutate(three_pct = round(three_total/state_total * 100 , 0)) %>%
  mutate(four_pct = round(four_total/state_total * 100, 0 )) 

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myRural_mapData <- left_join(myRural_mapData, fips_codes, by = c("state_fips" = "state_code")) 

myRural_mapData <- myRural_mapData %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region))

# --- maps --- #

myDifference_map <- myRural_mapData %>%
  select(region, one_pct, four_pct) %>%
  mutate(value = four_pct - one_pct) %>%
  filter(!is.na(region))

state_choropleth(myDifference_map,
  title = paste("Difference in % grants recieved by fourth quartile and first:", myStart, "-", myEnd),
  legend = "Difference in % recieved",
  zoom = continental_us_states,
  num_colors = 5) +
  scale_fill_brewer(palette="Purples", direction = -1) + 
  labs(fill = "Difference in % recieved")

}

myTime_pct()
myTime_pct(myEnd = 1981)
myTime_pct(myStart = 1981, myEnd = 2008)
myTime_pct(myStart = 2008)




```
The pre-Raegan years seemed to invest in rural communities the least! 

# Average Grant Size per Capita

I calculate the per capita amount of grant funding for each year and include years when the amount is zero. I find the average per capita amount of grant funding for each quartile. 

## Maps

```{r DATA-MANIPULATION-2, echo=FALSE, message=FALSE, warning= FALSE}

# --- data manipulation to get amount per person --- #

#getting region in here for the state_choropleth function
myWorking <- left_join(myWorking, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region))

# calculating grant award per capita for each county each year
myWorking <- myWorking %>%
  mutate(amount_perCap = if_else(!is.na(amount), amount / population,  0))

# -- first quartile maps -- #

# First quartile 
myPerCap_map1 <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)
  

state_choropleth(myPerCap_map1,
  title = "Average per capita grants amount given to first quartile of rural counties",
  legend = "Average per cap. $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr")+
  ggtitle(label = "Average per capita grants amount given to first quartile of rural counties", 
          subtitle =  "(Most urban counties)") + 
  labs(fill = "Average per cap. $")

#second quartile 
myPerCap_map2 <- myWorking %>%
  filter(rural_quarts == "2") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)
  

state_choropleth(myPerCap_map2,
  title = "Average per capita grants amount given to second quartile of rural counties",
  legend = "Average per cap. $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average per cap. $")

  
#third quartile 
myPerCap_map3 <- myWorking %>%
  filter(rural_quarts == "3") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)
  

state_choropleth(myPerCap_map3,
  title = "Average per capita grants amount given to thrid quartile of rural counties",
  legend = "Average per cap. $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average per cap. $")

#fourth quartile 
myPerCap_map4 <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)
  
state_choropleth(myPerCap_map4,
  title = "Average per capita grants amount given to fourth quartile of rural counties",
  legend = "Average per cap. $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average per cap. $")


# -- difference -- #


myPerCap_difference_map <- left_join(x = myPerCap_map4, y = myPerCap_map1, by = "region") %>%
  rename(value_4 = value.x) %>%
  rename(value_1 = value.y) %>%
  mutate(value = value_4 - value_1) %>%
  filter(!is.na(region))

state_choropleth(myPerCap_difference_map,
  title = "Difference in avg. per capita grant between fourth quartile and first",
  legend = "Difference in $",
  zoom = continental_us_states,
  num_colors = 7) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Difference in $")




```
Note: in North and South Dakota, the 3rd and 4th quartiles are both 100% rural, so I don't code any counties at the third quartile for rural in those states.

**Interpretation:** 

* On average in California, the counties in the first quartile for percentage of population that is in rural (ie the most urban counties) receive \$ 0.00 to \$0.14 of grant funding per person. 
* On average in California, the counties in the fourth quartile for % of population that is in rural (counties with the most rural) receive \$ 1.00 to \$4.79 of grant funding per person. 
* In California, counties in the fourth quartile receive \$ 0.82 - \$ 4.53 more grant funding per person than counties in the first quartile. 


In general, rural counties are receiving more per capita funding than urban.

## Table

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Getting columns together 
myPerCap_table <- left_join(myPerCap_map1, myPerCap_map2, by = "region") %>%
  rename(perCap_1 = value.x) %>%
  rename(perCap_2 = value.y) 
myPerCap_table <- left_join(myPerCap_table, myPerCap_map3, by = "region") %>%
  rename(perCap_3 = value) 
myPerCap_table <- left_join(myPerCap_table, myPerCap_map4, by = "region") %>%
  rename(perCap_4 = value) 
myPerCap_table <- left_join(myPerCap_table, myPerCap_difference_map, by = "region") %>%
  rename(perCap_diff = value) %>%
  select(-value_4, -value_1)

#get percent of rural for the state
state_pct_rural <- myWorking %>%
  select(region, rural_pct) %>%
  distinct() %>%
  group_by(region) %>%
  mutate(avg_rural_pct = mean(rural_pct)) %>%
  select(region, avg_rural_pct) %>%
  distinct()

#rounding two digets for all columns
myPerCap_table <- left_join(myPerCap_table, state_pct_rural, by = "region") %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  filter(region != "district of columbia") %>%
  filter(region != "puerto rico") 

myPerCap_table %>%
  arrange(-avg_rural_pct) %>%
  mutate(region = str_to_title(region)) %>%
  kbl(booktabs = T, 
      caption = "Avg. per capita grant ($)",
      col.names = c("State",
                    "First",
                    "Second", 
                    "Third",
                    "Fourth",
                    "Difference (four minus first)", 
                    "Avg % of county that's in rural")) %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Plot of avg % rural in state and difference between quarts 
ggplot(myPerCap_table, aes(x=avg_rural_pct, y=perCap_diff)) + geom_point() + geom_smooth(method = "lm") +theme_bw() + 
  xlab("Avg. % of rural people per County") +
  ylab("Difference in fouth and first quartile") + ggtitle("Avg. per capita spending")

```

When looking at the difference in the avg. per capita funding received, there is not an obvious trend with a counties' population that is in rural. This may be due to the amount of funding a state receives being highly correlated with population and so this final figure is not helpful for a per capita measure because grant funding is allocated to states in a way that tries to guarantee that per capita funding is equal across states.


## Time Differences: Per captia grant award ($)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

myTime_perCap <- function(myStart = 1965, myEnd = 2018){
myWorking <- myWorking %>%
    dplyr::filter(year >= myStart) %>%
    dplyr::filter(year < myEnd)
  
# First quartile 
myPerCap_map1 <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)

# Fourth quartile 
myPerCap_map4 <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(region) %>%
  mutate(mean_amount_perCap = mean(amount_perCap)) %>%
  select(region, mean_amount_perCap) %>%
  distinct() %>%
  rename(value = mean_amount_perCap)

# -- difference -- #
myPerCap_difference_map <- left_join(x = myPerCap_map4, y = myPerCap_map1, by = "region") %>%
  rename(value_4 = value.x) %>%
  rename(value_1 = value.y) %>%
  mutate(value = value_4 - value_1) %>%
  filter(!is.na(region))


state_choropleth(myPerCap_difference_map,
  title = paste("Difference in avg. per capita grant between fourth quartile and first", myStart, "-", myEnd),
  legend = "Difference in $",
  zoom = continental_us_states,
  num_colors = 7) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Difference in $")


}

myTime_perCap()
myTime_perCap(myEnd = 1981)
myTime_perCap(myStart = 1981, myEnd = 2008)
myTime_perCap(myStart = 2008)


```

It looks like California, Texas and the North East have all experienced a flip how they invest in rural communities.

** I'm trying think of way to organize this so that I more easily observe time trends then only looking at the maps**


# Average Grant Size (Total)

I find the average amount of grant funding for each quartile and include years when the amount is zero.

## Maps

```{r DATA-MANIPULATION-3, echo=FALSE, message=FALSE, warning= FALSE}

# --- data manipulation to get amount per person --- #

# calculating grant award for each county each year
myWorking <- myWorking %>%
  mutate(amount= if_else(!is.na(amount), amount,  0))

# -- first quartile maps -- #

# First quartile 
myAmount_map1 <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)
  

state_choropleth(myAmount_map1,
  title = "Average grants amount given to first quartile of rural counties",
  legend = "Average $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr")+
  ggtitle(label = "Average grants amount given to first quartile of rural counties", 
          subtitle =  "(most urban counties)") + 
  labs(fill = "Average $")

#second quartile 
myAmount_map2 <- myWorking %>%
  filter(rural_quarts == "2") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)
  

state_choropleth(myAmount_map2,
  title = "Average grants amount given to second quartile of rural counties",
  legend = "Average $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average $")

  
#third quartile 
myAmount_map3 <- myWorking %>%
  filter(rural_quarts == "3") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)
  

state_choropleth(myAmount_map3,
  title = "Average grants amount given to thrid quartile of rural counties",
  legend = "Average $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average $")

#fourth quartile 
myAmount_map4 <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)
  
state_choropleth(myAmount_map4,
  title = "Average grants amount given to fourth quartile of rural counties",
  legend = "Average $",
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr") + 
  labs(fill = "Average $")


# -- difference -- #


myAmount_difference_map <- left_join(x = myAmount_map4, y = myAmount_map1, by = "region") %>%
  rename(value_4 = value.x) %>%
  rename(value_1 = value.y) %>%
  mutate(value = value_4 - value_1) %>%
  filter(!is.na(region))

state_choropleth(myAmount_difference_map,
  title = "Difference in avg. grant between fourth quartile and first",
  legend = "Difference in $",
  zoom = continental_us_states,
  num_colors = 5) +
  scale_fill_brewer(palette="Purples", direction = -1) + 
  labs(fill = "Difference in $")




```


**Interpretation:** 

* On average in Florida, the counties in the first quartile for percentage of population that is in rural (ie the most urban counties) receive \$ 58K to \$107K of grant funding per person. 
* On average in Florida, the counties in the fourth quartile for % of population that is in rural (counties with the most rural) receive \$ 0 to \$3K of grant funding per person. 
* In Florida, counties in the fourth quartile receive \$ 109K - \$ 49K less grant funding per person than counties in the first quartile. 

## Table

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Getting columns together 
myAmount_table <- left_join(myAmount_map1, myAmount_map2, by = "region") %>%
  rename(perCap_1 = value.x) %>%
  rename(perCap_2 = value.y) 
myAmount_table <- left_join(myAmount_table, myAmount_map3, by = "region") %>%
  rename(perCap_3 = value) 
myAmount_table <- left_join(myAmount_table, myAmount_map4, by = "region") %>%
  rename(perCap_4 = value) 
myAmount_table <- left_join(myAmount_table, myAmount_difference_map, by = "region") %>%
  rename(perCap_diff = value) %>%
  select(-value_4, -value_1)

#get percent of rural for the state
state_pct_rural <- myWorking %>%
  select(region, rural_pct) %>%
  distinct() %>%
  group_by(region) %>%
  mutate(avg_rural_pct = mean(rural_pct)) %>%
  select(region, avg_rural_pct) %>%
  distinct()

#rounding two digets for all columns
myAmount_table <- left_join(myAmount_table, state_pct_rural, by = "region") %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  filter(region != "district of columbia") %>%
  filter(region != "puerto rico") 

myAmount_table %>%
  arrange(-avg_rural_pct) %>%
  mutate(region = str_to_title(region)) %>%
  kbl(booktabs = T, 
      caption = "Avg. grant ($)",
      col.names = c("State",
                    "First",
                    "Second", 
                    "Third",
                    "Fourth",
                    "Difference (four minus first)", 
                    "Avg % of county that's in rural")) %>%
  kable_styling(latex_options = c("scale_down","HOLD_position")) 


# Plot of avg % rural in state and difference between quarts 
ggplot(myAmount_table, aes(x=avg_rural_pct, y=perCap_diff)) + geom_point() + geom_smooth(method = "lm") +theme_bw() + 
  xlab("Avg. % of people in rural per County") +
  ylab("Difference in fouth and first quartile") + ggtitle("Avg. spending")

```

When looking at the difference in the avg. funding received, there is an upward trend with a counties' population that is in rural. 


## Time Differences: total grant award ($)

```{r, echo=FALSE, message=FALSE, warning=FALSE}

myTime_amount <- function(myStart = 1965, myEnd = 2018){
myWorking <- myWorking %>%
    dplyr::filter(year >= myStart) %>%
    dplyr::filter(year < myEnd)
  
# First quartile 
myAmount_map1 <- myWorking %>%
  filter(rural_quarts == "1") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)

# Fourth quartile 
myAmount_map4 <- myWorking %>%
  filter(rural_quarts == "4") %>%
  group_by(region) %>%
  mutate(mean_amount = mean(amount)) %>%
  select(region, mean_amount) %>%
  distinct() %>%
  rename(value = mean_amount)

# -- difference -- #
myAmount_difference_map <- left_join(x = myAmount_map4, y = myAmount_map1, by = "region") %>%
  rename(value_4 = value.x) %>%
  rename(value_1 = value.y) %>%
  mutate(value = value_4 - value_1) %>%
  filter(!is.na(region))


state_choropleth(myAmount_difference_map,
  title = paste("Difference in avg. grant between fourth quartile and first", myStart, "-", myEnd),
  legend = "Difference in $",
  zoom = continental_us_states,
  num_colors = 5) +
  scale_fill_brewer(palette="Purples", direction = -1) + 
  labs(fill = "Difference in $")


}

myTime_amount()
myTime_amount(myEnd = 1981)
myTime_amount(myStart = 1981, myEnd = 2008)
myTime_amount(myStart = 2008)


```

No major time trends are jumping out to me! 

