---
title: "Quartiles of Grants Described: Land Aquisitions"
author: "Andie Creel"
date: "7/21/2021"
output: 
  pdf_document:
    number_sections: true
    extra_dependencies: ["float"]

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'Exploratory_Output/descriptive/grant_quarts_acquisitions.pdf')) })
---

**This document only looks at grants that involved land acquisition (grant type is either A for acquisition or C for combination)**

Brown colors tend to indicated negative numbers and purple tend to indicate positive numbers. 

Active years: 1975-2018 because the non-hispanic white characteristic is not available prior to 1980


```{r, include=FALSE}
#R Markdown stuff
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
options(tinytex.verbose = TRUE)
```

```{r setup, include=FALSE}
library(vroom)
library(tidyr)
library(stringr)
library(usmap)
library(tidycensus)
library(ggplot2)
library(fixest)
library(kableExtra)
library(gtools)
library(forcats)
library("wesanderson")
options(scipen=99999)
library(janitor)
library(dplyr)
# ---- #
library(choroplethr)
library(choroplethrMaps)
library(usdata)
library(maps)
data(fips_codes)
data(continental_us_states)
# --- #

myWorking <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv")

# additional data cleaning ------------

# Removing counties whose FIPSs code I didn't identify and don't have demo data for (1975-2018)
myWorking <- myWorking %>%
  filter(!is.na(fips)) %>%
  distinct() %>%
  mutate(type = str_sub(type,0,1)) %>%
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>%
  filter(!is.na(white_pct)) %>%
  dplyr::mutate(poc_pct = 100 - white_pct)


myWorking <- myWorking %>%
  filter(type != "P") %>% # Planning 
  filter(type != "D") %>% # Development
  filter(type != "R") # redevelopment 

# Grouping outcome variables by decade (number of grants received, amount received, average amount received per capita per decade, avg # grants per 100,000 people per decade)
myWorking_decade <- myWorking %>%
  group_by(fips, merge_year) %>%
  mutate(decade_grants = sum(got_grant)) %>%
  mutate(decade_amount = sum(amount, na.rm = TRUE)) %>%
  mutate(decade_avg_per_cap_amount = sum(amount, na.rm = TRUE)/mean(annual_population, na.rm = TRUE)) %>%
  mutate(decade_avg_per_cap_grants = sum(got_grant)/mean((annual_population/100000), na.rm = TRUE)) %>%
  select(state_fips, ends_with("_pct"), starts_with("decade_")) %>%
  ungroup() %>%
  distinct()

#getting average POC per county
state_pct_POC <- myWorking_decade %>%
  select(state_fips, poc_pct) %>%
  distinct() %>%
  group_by(state_fips) %>%
  mutate(avg_poc_pct = mean(poc_pct)) %>%
  select(state_fips, avg_poc_pct) %>%
  distinct()

myWorking_decade <- left_join(myWorking_decade, state_pct_POC, by = "state_fips" )

# ------- Get quartiles of grants received for each decade ------- #
myQuants <- myWorking_decade %>%
  dplyr::group_by(state_fips, merge_year) %>%
  dplyr::summarise(quants = list(quantile(decade_grants, probs = c(.25, .5, .75)))) %>%
  unnest_wider(quants) %>%
  dplyr::rename(decade_grants_25 = `25%`) %>%
  dplyr::rename(decade_grants_50 = `50%`) %>%
  dplyr::rename(decade_grants_75 = `75%`)

#joining back in
myWorking_decade <- left_join(myWorking_decade, myQuants, by = c("state_fips", "merge_year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking_decade <- myWorking_decade %>%
  dplyr::mutate(decade_grants_quants = if_else(decade_grants > decade_grants_75, "4", "")) %>% # fourth quartile 
  dplyr::mutate(decade_grants_quants = if_else(decade_grants <=decade_grants_75, "3", decade_grants_quants)) %>% # third quartile
  dplyr::mutate(decade_grants_quants = if_else(decade_grants <= decade_grants_50, "2", decade_grants_quants)) %>% # second quartile
  dplyr::mutate(decade_grants_quants = if_else(decade_grants <= decade_grants_25, "1", decade_grants_quants)) %>% # first quartile 
  dplyr::select(-decade_grants_75, -decade_grants_50, - decade_grants_25) %>%
  dplyr::mutate(decade_grants_quants = as.factor(decade_grants_quants))

# ------- Get quartiles of amount of grant funding for each decade ------- #
myQuants <- myWorking_decade %>%
  dplyr::group_by(state_fips, merge_year) %>%
  dplyr::summarise(quants = list(quantile(decade_amount, probs = c(.25, .5, .75)))) %>%
  unnest_wider(quants) %>%
  dplyr::rename(decade_amount_25 = `25%`) %>%
  dplyr::rename(decade_amount_50 = `50%`) %>%
  dplyr::rename(decade_amount_75 = `75%`)

#joining back in
myWorking_decade <- left_join(myWorking_decade, myQuants, by = c("state_fips", "merge_year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking_decade <- myWorking_decade %>%
  dplyr::mutate(decade_amount_quants = if_else(decade_amount > decade_amount_75, "4", "")) %>% # fourth quartile 
  dplyr::mutate(decade_amount_quants = if_else(decade_amount <=decade_amount_75, "3", decade_amount_quants)) %>% # third quartile
  dplyr::mutate(decade_amount_quants = if_else(decade_amount <= decade_amount_50, "2", decade_amount_quants)) %>% # second quartile
  dplyr::mutate(decade_amount_quants = if_else(decade_amount <= decade_amount_25, "1", decade_amount_quants)) %>% # first quartile 
  dplyr::select(-decade_amount_75, -decade_amount_50, - decade_amount_25) %>%
  dplyr::mutate(decade_amount_quants = as.factor(decade_amount_quants))

# ------- Get quartiles of amount of grant funding PER CAPITA for each decade ------- #
myQuants <- myWorking_decade %>%
  dplyr::group_by(state_fips, merge_year) %>%
  dplyr::summarise(quants = list(quantile(decade_avg_per_cap_amount, probs = c(.25, .5, .75), na.rm = TRUE))) %>%
  unnest_wider(quants) %>%
  dplyr::rename(decade_avg_per_cap_amount_25 = `25%`) %>%
  dplyr::rename(decade_avg_per_cap_amount_50 = `50%`) %>%
  dplyr::rename(decade_avg_per_cap_amount_75 = `75%`)

#joining back in
myWorking_decade <- left_join(myWorking_decade, myQuants, by = c("state_fips", "merge_year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking_decade <- myWorking_decade %>%
  dplyr::mutate(decade_avg_per_cap_amount_quants = if_else(decade_avg_per_cap_amount > decade_avg_per_cap_amount_75, "4", "")) %>% # fourth quartile 
  dplyr::mutate(decade_avg_per_cap_amount_quants = if_else(decade_avg_per_cap_amount <=decade_avg_per_cap_amount_75, "3", decade_avg_per_cap_amount_quants)) %>% # third quartile
  dplyr::mutate(decade_avg_per_cap_amount_quants = if_else(decade_avg_per_cap_amount <= decade_avg_per_cap_amount_50, "2", decade_avg_per_cap_amount_quants)) %>% # second quartile
  dplyr::mutate(decade_avg_per_cap_amount_quants = if_else(decade_avg_per_cap_amount <= decade_avg_per_cap_amount_25, "1", decade_avg_per_cap_amount_quants)) %>% # first quartile 
  dplyr::select(-decade_avg_per_cap_amount_75, -decade_avg_per_cap_amount_50, - decade_avg_per_cap_amount_25) %>%
  dplyr::mutate(decade_avg_per_cap_amount_quants = as.factor(decade_avg_per_cap_amount_quants))

# ------- Get quartiles of number of grants  PER CAPITA for each decade ------- #
myQuants <- myWorking_decade %>%
  dplyr::group_by(state_fips, merge_year) %>%
  dplyr::summarise(quants = list(quantile(decade_avg_per_cap_grants, probs = c(.25, .5, .75), na.rm = TRUE))) %>%
  unnest_wider(quants) %>%
  dplyr::rename(decade_avg_per_cap_grants_25 = `25%`) %>%
  dplyr::rename(decade_avg_per_cap_grants_50 = `50%`) %>%
  dplyr::rename(decade_avg_per_cap_grants_75 = `75%`)

#joining back in
myWorking_decade <- left_join(myWorking_decade, myQuants, by = c("state_fips", "merge_year"))
rm(myQuants)

#making a factor variable for the quartiles (by state and year)
myWorking_decade <- myWorking_decade %>%
  dplyr::mutate(decade_avg_per_cap_grants_quants = if_else(decade_avg_per_cap_grants > decade_avg_per_cap_grants_75, "4", "")) %>% # fourth quartile 
  dplyr::mutate(decade_avg_per_cap_grants_quants = if_else(decade_avg_per_cap_grants <=decade_avg_per_cap_grants_75, "3", decade_avg_per_cap_grants_quants)) %>% # third quartile
  dplyr::mutate(decade_avg_per_cap_grants_quants = if_else(decade_avg_per_cap_grants <= decade_avg_per_cap_grants_50, "2", decade_avg_per_cap_grants_quants)) %>% # second quartile
  dplyr::mutate(decade_avg_per_cap_grants_quants = if_else(decade_avg_per_cap_grants <= decade_avg_per_cap_grants_25, "1", decade_avg_per_cap_grants_quants)) %>% # first quartile 
  dplyr::select(-decade_avg_per_cap_grants_75, -decade_avg_per_cap_grants_50, - decade_avg_per_cap_grants_25) %>%
  dplyr::mutate(decade_avg_per_cap_grants_quants = as.factor(decade_avg_per_cap_grants_quants))

```

# Maps

## Quantity of grants per capita 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# QUANTITY OF GRANTS PER CAPITA

# fourth quartile is receiving most amount of grants 

# difference in % POC from 1st quart and 4th quart -------------

myMap_poc <- myWorking_decade %>%
  select(state_fips, fips, merge_year, poc_pct, decade_avg_per_cap_grants_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_grants_quants, values_from = c(poc_pct)) %>%
  group_by(state_fips) %>%
  mutate(poc_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(poc_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(poc_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(poc_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("poc_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_poc <- left_join(myMap_poc, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(poc_pct_4 - poc_pct_1), 2)

state_choropleth(myMap_poc,
  zoom = continental_us_states) +
  # scale_fill_brewer(palette="Purples")+
  scale_fill_brewer(palette="YlOrBr", direction = -1) +
  # scale_fill_brewer(palette="PuOr") +
  ggtitle(label = "POC: Top quartile for # grants per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % of POC (top minus bottom)") + 
  labs(fill = "Difference in %") +
  theme(legend.position="bottom")


# difference in % poverty from 1st quart and 4th quart -------------

myMap_pov <- myWorking_decade %>%
  select(state_fips, fips, merge_year, inc_below_pov_pct, decade_avg_per_cap_grants_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_grants_quants, values_from = c(inc_below_pov_pct)) %>%
  group_by(state_fips) %>%
  mutate(inc_below_pov_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("inc_below_pov_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_pov <- left_join(myMap_pov, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(inc_below_pov_pct_4 - inc_below_pov_pct_1), 2)

state_choropleth(myMap_pov,
  zoom = continental_us_states) +
  # scale_fill_brewer(palette="YlOrBr", direction = -1)+
  scale_fill_brewer(palette="Purples")+
  ggtitle(label = "Poverty: Top quartile for # grants per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % in poverty (top minus bottom)") + 
  labs(fill = "Difference in % ") +
  theme(legend.position="bottom")


# difference in % urban from 1st quart and 4th quart -------------

myMap_rural <- myWorking_decade %>%
  select(state_fips, fips, merge_year, rural_pct, decade_avg_per_cap_grants_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_grants_quants, values_from = c(rural_pct)) %>%
  group_by(state_fips) %>%
  mutate(rural_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(rural_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(rural_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(rural_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("rural_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_rural <- left_join(myMap_rural, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(rural_pct_4 - rural_pct_1), 2)

state_choropleth(myMap_rural,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="Purples")+
  ggtitle(label = "Rural: Top quartile for # grants per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % rural (top minus bottom)") + 
  labs(fill = "Difference in %") +
  theme(legend.position="bottom")

```
Interpretation: 
In CA, the counties that receive the most number grants per capita in a given decade have **10-23% less POC** than counties that receive the least number of grants per capita.

North east, CA, IL, CO all have strong neg correlation with quantity per cap and POC. 

In GA, the counties that receive the most number grants per capita in a given decade have **8-9% more people in poverty** than counties that receive the least number of grants per capita.

No major regeion or national trends with quantity per cap and poverty. 

In CO, the counties that receive the most number grants per capita in a given decade have **52-62% more rural people** than counties that receive the least number of grants per capita.

There is a major correlation betwen rural and quanityt per cap.


## Amount of grant funding per capita 


```{r, echo=FALSE, message=FALSE, warning= FALSE}
rm(myMap_poc, myMap_pov, myMap_rural)

# AMOUNT OF GRANT $$ PER CAPITA

# difference in % POC from 1st quart and 4th quart -------------

myMap_poc <- myWorking_decade %>%
  select(state_fips, fips, merge_year, poc_pct, decade_avg_per_cap_amount_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_amount_quants, values_from = c(poc_pct)) %>%
  group_by(state_fips) %>%
  mutate(poc_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(poc_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(poc_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(poc_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("poc_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_poc <- left_join(myMap_poc, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(poc_pct_4 - poc_pct_1), 2)

state_choropleth(myMap_poc,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr", direction = -1)+
  # scale_fill_brewer(palette="PuOr")+
  ggtitle(label = "POC: Top quartile for amount of funding per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % of POC (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


 # difference in % poverty from 1st quart and 4th quart -------------

myMap_pov <- myWorking_decade %>%
  select(state_fips, fips, merge_year, inc_below_pov_pct, decade_avg_per_cap_amount_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_amount_quants, values_from = c(inc_below_pov_pct)) %>%
  group_by(state_fips) %>%
  mutate(inc_below_pov_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("inc_below_pov_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_pov <- left_join(myMap_pov, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(inc_below_pov_pct_4 - inc_below_pov_pct_1), 2)

state_choropleth(myMap_pov,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="PuOr")+
  # scale_fill_brewer(palette="Purples")+
  ggtitle(label = "Poverty: Top quartile for amount of funding per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % in poverty (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


# difference in % urban from 1st quart and 4th quart -------------

myMap_rural <- myWorking_decade %>%
  select(state_fips, fips, merge_year, rural_pct, decade_avg_per_cap_amount_quants) %>%
  pivot_wider(names_from = decade_avg_per_cap_amount_quants, values_from = c(rural_pct)) %>%
  group_by(state_fips) %>%
  mutate(rural_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(rural_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(rural_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(rural_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("rural_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_rural <- left_join(myMap_rural, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(rural_pct_4 - rural_pct_1), 2)

state_choropleth(myMap_rural,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="Purples")+
  # scale_fill_brewer(palette="PuOr")+
  ggtitle(label = "Rural: Top quartile for amount of funding per capita vs the bottom quartile", 
          subtitle =  "Difference in avg % rural (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


```
Interpretation: 
In CA, the counties that receive the most amount of grant funding per capita in a given decade have **7-21% less POC** than counties that receive the least amount of grant funding per capita.

There are several states that with a moderate to strong negative correlations between POC and funding per cap

In NM, the counties that receive the most amount of grant funding per capita in a given decade have **1-8% less people in poverty** than counties that receive the least amount of grant funding per capita.

No major trend

In almost every state, the counties that receive the most most amount of grant funding per capita in a given decade have **more rural people** than counties that receive the least amount of grant funding per capita.

Very strong positive correlation between rural and funding per cap

## Quantity of grants


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# QUANTITY OF GRANTS 

# fourth quartile is receiving most amount of grants 

# difference in % POC from 1st quart and 4th quart -------------

myMap_poc <- myWorking_decade %>%
  select(state_fips, fips, merge_year, poc_pct, decade_grants_quants) %>%
  pivot_wider(names_from = decade_grants_quants, values_from = c(poc_pct)) %>%
  group_by(state_fips) %>%
  mutate(poc_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(poc_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(poc_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(poc_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("poc_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_poc <- left_join(myMap_poc, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(poc_pct_4 - poc_pct_1), 2)

state_choropleth(myMap_poc,
  zoom = continental_us_states) +
  # scale_fill_brewer(palette="YlOrBr", direction = -1)+
  scale_fill_brewer(palette="PuOr")+
  ggtitle(label = "POC: Top quartile for # grants vs the bottom quartile", 
          subtitle =  "Difference in avg % of POC (top minus bottom)") + 
  labs(fill = "Difference in %") +
  theme(legend.position="bottom")


# difference in % poverty from 1st quart and 4th quart -------------

myMap_pov <- myWorking_decade %>%
  select(state_fips, fips, merge_year, inc_below_pov_pct, decade_grants_quants) %>%
  pivot_wider(names_from = decade_grants_quants, values_from = c(inc_below_pov_pct)) %>%
  group_by(state_fips) %>%
  mutate(inc_below_pov_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("inc_below_pov_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_pov <- left_join(myMap_pov, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(inc_below_pov_pct_4 - inc_below_pov_pct_1), 2)

state_choropleth(myMap_pov,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr", direction = -1)+
  # scale_fill_brewer(palette="PuOr")+
  ggtitle(label = "Poverty: Top quartile for # grants vs the bottom quartile", 
          subtitle =  "Difference in avg % in poverty (top minus bottom)") + 
  labs(fill = "Difference in % ") +
  theme(legend.position="bottom")


# difference in % urban from 1st quart and 4th quart -------------

myMap_rural <- myWorking_decade %>%
  select(state_fips, fips, merge_year, rural_pct, decade_grants_quants) %>%
  pivot_wider(names_from = decade_grants_quants, values_from = c(rural_pct)) %>%
  group_by(state_fips) %>%
  mutate(rural_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(rural_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(rural_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(rural_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("rural_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_rural <- left_join(myMap_rural, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(rural_pct_4 - rural_pct_1), 2)

state_choropleth(myMap_rural,
  zoom = continental_us_states) +
  # scale_fill_brewer(palette="PuOr")+
  scale_fill_brewer(palette="YlOrBr", direction = -1)+
  ggtitle(label = "Rural: Top quartile for # grants vs the bottom quartile", 
          subtitle =  "Difference in avg % rural (top minus bottom)") + 
  labs(fill = "Difference in %") +
  theme(legend.position="bottom")


```
Interpretation: 
In MD and NJ and other, the counties that receive the highest quantity of grants in a given decade have **8-18% less POC** than counties that receive the smallest quantity of grants.

This is a moderate to strong trend in some states.

In ND, the counties that receive the highest quantity of grants in a given decade have **5-10% less people in poverty** than counties that receive the smallest quantity of grants.

There is a weak trend

For the majority of the country, the counties that receive the highest quantity of grants in a given decade have **are less rural** than counties that receive the smallest quantity of grants.

There is a very strong trends


## Amount of grant funding


```{r, echo=FALSE, message=FALSE, warning= FALSE}
rm(myMap_poc, myMap_pov, myMap_rural)

# AMOUNT OF GRANT $$

# difference in % POC from 1st quart and 4th quart -------------

myMap_poc <- myWorking_decade %>%
  select(state_fips, fips, merge_year, poc_pct, decade_amount_quants) %>%
  pivot_wider(names_from = decade_amount_quants, values_from = c(poc_pct)) %>%
  group_by(state_fips) %>%
  mutate(poc_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(poc_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(poc_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(poc_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("poc_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_poc <- left_join(myMap_poc, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(poc_pct_4 - poc_pct_1), 2)

state_choropleth(myMap_poc,
  zoom = continental_us_states) +
  # scale_fill_brewer(palette="YlOrBr", direction = -1)+
  scale_fill_brewer(palette="PuOr")+
  ggtitle(label = "POC: Top quartile for amount of funding vs the bottom quartile", 
          subtitle =  "Difference in avg % of POC (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


 # difference in % poverty from 1st quart and 4th quart -------------

myMap_pov <- myWorking_decade %>%
  select(state_fips, fips, merge_year, inc_below_pov_pct, decade_amount_quants) %>%
  pivot_wider(names_from = decade_amount_quants, values_from = c(inc_below_pov_pct)) %>%
  group_by(state_fips) %>%
  mutate(inc_below_pov_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(inc_below_pov_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("inc_below_pov_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_pov <- left_join(myMap_pov, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(inc_below_pov_pct_4 - inc_below_pov_pct_1), 2)

state_choropleth(myMap_pov,
  zoom = continental_us_states,
  num_colors = 7) +
  scale_fill_brewer(palette="PuOr")+
  # scale_fill_brewer(palette="YlOrBr", direction = -1)+
  # scale_fill_brewer(palette="Purples")+
  ggtitle(label = "Poverty: Top quartile for amount of funding vs the bottom quartile", 
          subtitle =  "Difference in avg % in poverty (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


# difference in % urban from 1st quart and 4th quart -------------

myMap_rural <- myWorking_decade %>%
  select(state_fips, fips, merge_year, rural_pct, decade_amount_quants) %>%
  pivot_wider(names_from = decade_amount_quants, values_from = c(rural_pct)) %>%
  group_by(state_fips) %>%
  mutate(rural_pct_4 = mean(`4`, na.rm = TRUE)) %>%
  mutate(rural_pct_3 = mean(`3`, na.rm = TRUE)) %>%
  mutate(rural_pct_2 = mean(`2`, na.rm = TRUE)) %>%
  mutate(rural_pct_1 = mean(`1`, na.rm = TRUE)) %>%
  select(state_fips, starts_with("rural_pct_")) %>%
  distinct()

#getting state name for choropleth maps 
fips_codes <- fips_codes %>%
  select(state_code, state_name) %>%
  distinct()

myMap_rural <- left_join(myMap_rural, fips_codes, by = c("state_fips" = "state_code")) %>% 
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  mutate(value = round(rural_pct_4 - rural_pct_1), 2)

state_choropleth(myMap_rural,
  zoom = continental_us_states) +
  scale_fill_brewer(palette="YlOrBr", direction = -1)+
  # scale_fill_brewer(palette="Purples")+
  ggtitle(label = "Rural: Top quartile for amount of funding vs the bottom quartile", 
          subtitle =  "Difference in avg % rural (top minus bottom)") + 
  labs(fill = "Difference in % ")+
  theme(legend.position="bottom")


```
In DE, the counties that receive the highest amount of grant funding in a given decade have **4-27% less POC** than counties that receive the smallest amount of grant funding.

There are some states with a moderate trend.

In AZ, the counties that receive the highest amount of grant funding in a given decade have **4-11% less people in poverty** than counties that receive the lowest amount of grant funding.

There are some states with a moderate trend.

For the majority of the county, the counties that receive the highest amount of grant fundin in a given decade have **are less rural** than counties that receive the smallest amount of grant funding.

There is a very strong national trend
