---
title: "Suits Index"
author: "Andie Creel"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE, message=FALSE}
library(vroom)
library(tidyr)
library(stringr)
library(ggplot2)
library(gtools)
library(forcats)
options(scipen=99999)
library(janitor)
library(dplyr)
library(zoo)
library(runner)
source('09.functions.R')
```

Thanks to Matt Ashenfarb for providing helpful code on the suites index. 
```{r, include=FALSE, message=FALSE}
#-------------------------------------------------------------------------------
# Basic data cleaning
# Grant amounts are already adjusted for inflation, adjusting median income (2018 dollars)
#-------------------------------------------------------------------------------

# made in 01.data_wrangle.R
myWorking <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv") 

#Basic data cleaning
myWorking <- myWorking %>%
  mutate(type = str_sub(type,0,1)) %>% #getting rid of a space 
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>% # getting state fips code
  filter(!is.na(white_pct)) %>% # getting rid of rows where I couldn't get their demographic data
  dplyr::mutate(poc_pct = 100 - white_pct) %>% # calculating % people of color as non-hispanic white
  mutate(amount = if_else(is.na(amount), 0, amount)) %>%  # adding 0s for amount of grant rewarded
  filter(type != "P") # not including planning grants bc they're given to states, not counties

# working years: 1965-2018
myWorking %>%
  select(real_year) %>%
  distinct() %>%
  summary()

```

# Set Up and Definitions 

The **outcomes of interest** is how the amount of LWCF funds and quantity of grants are distributed across demographic characteristics of interest. 

**Median household income** for a county is the demographic characteristic focused on in this write up. I use the average median houshold income (averaged from 1965-2018 for each county). 

I use **two different per capita measurements** for the amount of LWCF funds and quantity of grants:

1. Funds (or quantity of grants) given to a county  divided by the population of the county at the time of LWCF distribution. This captures how the program was administered through time
2. The cumulative amount of funds (or quantity of grants) given to a county from 1965-2018 divided by the population of the county in 2020 (estimates for population of each county is not available for each county in 2018 in the American Community Survey). This captures how individuals today are effected by the LWCF program.

I use per capita metrics for amount of funding and quantity of grants to control for a county's population. 

```{r, include=FALSE}
# Need to get two different per capital award amounts: per cap on nearest year to award amount (how it's been distributed historically) and per cap on most recent year (snap shot on how current people have been invested in)

#Steps: 
#1) Set up the two different per caps 
#1.5) Set up median income on X axis instead of % race (maybe write function where we can say variable on X, and give it the dataset (ex, just MTs dataset) so that this can be neater and quicker)
#2) National Suits for each (line graph)
#3) Suits for each state (choropleth)

#-------------------------------------------------------------------------------
# First per cap: per cap through time aka how it's been implemented through time
#-------------------------------------------------------------------------------
myWorking_per_cap_1 <- myWorking %>%
    mutate(amount_per_cap = amount / population) %>% # amount per capita each year
    mutate(quantity_per_cap = got_grant/ population) %>% #  quantity per capita each year
    group_by(fips) %>% # group by county
    mutate(amount_per_cap_cum = sum(amount_per_cap)) %>% # cumulative amount per cap
    mutate(quantity_per_cap_cum = sum(quantity_per_cap)) %>% # cumulative quantity per cap
    mutate(avg_med_income_house = mean(med_income_house, na.rm = TRUE)) %>% #average median income
    select(state_fips, fips, amount_per_cap_cum, quantity_per_cap_cum, avg_med_income_house) %>% # set up reduce down to one entry per county
    distinct() # reduce


#-------------------------------------------------------------------------------
#Second per cap: per current population aka snapshot of how people today are affect 
#-------------------------------------------------------------------------------

myLatest_pop <- read.csv("Datasets/raw_data/co-est2021-alldata.csv") %>%
  select(STATE, COUNTY, POPESTIMATE2020) %>%
  mutate(STATE = str_pad(STATE, width = 2, pad ="0")) %>%
  mutate(COUNTY = str_pad(COUNTY, width = 3, pad = 0)) %>%
  mutate(fips = paste0(STATE, COUNTY)) %>%
  rename(pop_2020 = POPESTIMATE2020) %>%
  select(fips, pop_2020)

myWorking_per_cap_2.a <- myWorking %>%
    group_by(fips) %>% # group by county
    mutate(amount_cum = sum(amount)) %>% # cumulative amount per cap
    mutate(quantity_cum = sum(got_grant)) %>% # cumulative quantity per cap
    mutate(avg_med_income_house = mean(med_income_house, na.rm = TRUE)) %>% #average median income
    select(state_fips, fips, amount_cum, quantity_cum, avg_med_income_house) %>% # set up reduce down to one entry per county
    distinct() # reduce

myWorking_per_cap_2 <- left_join(myWorking_per_cap_2.a, myLatest_pop, "fips") %>%
   mutate(amount_per_cap_cum = amount_cum / pop_2020) %>% # amount per capita in 2020
    mutate(quantity_per_cap_cum = quantity_cum / pop_2020)  #  quantity per capita in 2020
rm(myWorking_per_cap_2.a, myLatest_pop)  


```

# Nation Picture

### First per capita 
Funds (or quantity of grants) given to a county  divided by the population of the county at the time of LWCF distribution. This captures how the program was administered through time

```{r}

#-------------------------------------------------------------------------------
# First per cap: results for nation 
#-------------------------------------------------------------------------------

myResult_1 <- getSuits(myWorking_per_cap_1)
printSuits(myResult_1)
```

### Second per capita 
The cumulative amount of funds (or quantity of grants) given to a county from 1965-2018 divided by the population of the county in 2020. This captures how individuals today are effected by the LWCF program.

```{r}

#-------------------------------------------------------------------------------
# First per cap: results for nation 
#-------------------------------------------------------------------------------

myResult_2 <- getSuits(myWorking_per_cap_2)
printSuits(myResult_2)
```


# State by State Picture






<!-- # Ex-Confederate States -->
<!-- Code isn't done yet. -->
```{r, include= FALSE}
myWorking_cum_south <- myWorking %>%
    filter(type != "P") %>% # not including planning grants bc they're given to states, not counties
    group_by(fips) %>%
    mutate(quantity = sum(got_grant)) %>% #grants through all time
    mutate(amount = sum(amount, na.rm = TRUE)) %>% # amoutn through all time
    mutate(avg_white_ppl= mean(white)) %>%
    select(fips, state_fips, quantity, quantity, amount, avg_white_ppl) %>%
    ungroup() %>%
    distinct() %>%
    filter(state_fips =='48' | state_fips =='05'| state_fips == '22' | state_fips == '47' | state_fips =='28' | state_fips == '01' | state_fips == '13' | state_fips == '12' | state_fips == '45' | state_fips == '37' | state_fips == '51')
    # confederacy: Texas, Arkansas, Louisiana, Tennessee, Mississippi, Alabama, Georgia, Florida, South Carolina, North Carolina and Virginia

```
