---
title: "Data Summary Stats"
author: "Andie Creel"
date: "2023-01-29"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(vroom)
library(tidyr)
library(stringr)
library(ggplot2)
library(gtools)
library(forcats)
options(scipen=99999)
library(janitor)
library(dplyr)
library(zoo)
library(runner)
source('09.functions.R')
library(fixest)
library(kableExtra)
library(stargazer)

# ----maps ---- #
library(choroplethr)
library(choroplethrMaps)
library(usdata)
library(maps)
library(tidycensus)
data("fips_codes")
data("state.regions")
library(RColorBrewer)
```

## Summary Statistics about our dataset 

### How many individual grants did TWS have data on?

```{r, results='asis', message=FALSE}
# made in 01.data_wrangle.R
myTWS <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv") 

myTWS <- myTWS %>% 
  select(fips, year, type, amount) %>%
  filter(!is.na(type)) %>%
  filter(amount != 0) %>%
  distinct()

tws_grant_num <- length(myTWS$type)


type_percents <- count(myTWS, type) %>%
  mutate(perc = round(n/tws_grant_num, 2))

type_percents
```

TWS had data on `r tws_grant_num` grants where they knew the grant type, the amount of funding reward, and the county that the project was located inside of.

Types of LWCF projects 

* A: Land Acquisition
* D: Development 
* R: Redevelopment 
* C: Combination
* P: Planning

### How many individual grants stay in our dataset (where we know median income, poc % and rural %)

```{r, results='asis', message=FALSE}
# made in 01.data_wrangle.R
myWorking <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv") 

#cleaning to relevant variables 
myWorking_clean <- myWorking %>%
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>% # getting state fips code
  filter(!is.na(white_pct)) %>% # getting rid of rows where I couldn't get their demographic data
  dplyr::mutate(poc_pct = 100 - white_pct) %>% # calculating % people of color as non-hispanic white
  mutate(amount = if_else(is.na(amount), 0, amount)) %>% # adding 0s for amount of grant rewarded
  select(state_fips, fips, real_year, merge_year, annual_population, amount, poc_pct, med_income_house, rural_pct, type) %>%
  filter(amount != 0) %>% # dropping grants where we don't know the grant amount
  # not dropping grants where we don't know med income bc that would drop a decades worth of data (dont have med income in 1970 census)
  filter(!is.na(poc_pct)) %>% #dropping grants where we don't know POC
  filter(!is.na(rural_pct)) #dropping grants where we don't know rural 

my_grant_num <- length(myWorking_clean$state_fips)


```

Our data set has `r my_grant_num` grants. 



### Summary Table

```{r, results='asis'}
makeSummaryTable <- function(df){
  
  # getting started with type break down
  mySumTable <- count(df, type) %>%
    mutate(perc = round(n/length(df$type), 2))
  
  # getting average grant size by type
  myAvg_grant_size <-  df %>%
    group_by(type) %>%
    mutate(avg_amount = mean(amount)) %>%
    select(type, avg_amount) %>%
    distinct()
    
  mySumTable <- left_join(mySumTable, myAvg_grant_size, by = "type")
  
  #creating row for total
  mySumTable[5, 1] <- "n"
  mySumTable[5, 2] <- sum(mySumTable$n, na.rm = T)
  mySumTable[5, 3] <- sum(mySumTable$perc, na.rm = T)
  mySumTable[5, 4] <- mean(df$amount)
  
  # cleaning up table 
  mySumTable <- mySumTable %>%
    mutate(avg_amount = round(avg_amount, 2)) %>%
    mutate(perc = perc*100)
  
  # replacing type 
  mySumTable$type[1] <- "Acquisition"
  mySumTable$type[2] <- "Combination"
  mySumTable$type[3] <- "Development"
  mySumTable$type[4] <- "Redevelopment"
  
  # renaming column names 
  colnames(mySumTable) <- c("Type", "Count", "Perc.", "Avg. Amount (2018$)")
  
  return(mySumTable)

}

mySumTable <- makeSummaryTable(myWorking_clean)

stargazer(mySumTable, type = "latex", summary = F, rownames = F, title = "Summary of grants included in our dataset")

mySumTable

```

### what is the average investment per capita ? 


```{r}
#restarting this every time to keep clear what I'm doing. 
rm(list = ls())

# -----------------------------------------------------------------------------
# Dataset on  -- poc and rural
# -----------------------------------------------------------------------------

# made in 01.data_wrangle.R
myWorking <- vroom("Datasets/clean_data/lwcf_annualPop_deccenialDemographics.csv") 

# cleaning to relevant variables 
myWorking_clean <- myWorking %>%
  mutate(state_fips = as.factor(str_sub(fips, 1, 2))) %>% # getting state fips code
  filter(!is.na(white_pct)) %>% # getting rid of rows where I couldn't get their demographic data
  dplyr::mutate(poc_pct = 100 - white_pct) %>% # calculating % people of color as non-hispanic white
  mutate(amount = if_else(is.na(amount), 0, amount)) %>% # adding 0s for amount of grant rewarded
  select(state_fips, fips, real_year, merge_year, annual_population, amount, poc_pct, med_income_house, rural_pct) %>%
  filter(amount != 0) %>% # dropping grants where we don't know the grant amount
  # not dropping grants where we don't know med income bc that would drop a decades worth of data (dont have med income in 1970 census)
  filter(!is.na(poc_pct)) %>% #dropping grants where we don't know POC
  filter(!is.na(rural_pct)) #dropping grants where we don't know rural 

# creating new variable where amount is defined as total amount a county got in a single year
myWorking_fips_year <- myWorking_clean %>%
  group_by(fips, real_year) %>%
  mutate(amount = sum(amount)) %>% #total amount a county got in a single year
  ungroup() %>%
  distinct()

# per cap: per cap through time aka how it's been implemented through time
myWorking_per_cap <- myWorking_fips_year %>%
    mutate(amount_per_cap = amount / annual_population) %>% # amount per capita each year
    group_by(fips) %>% # group by county
    mutate(amount_per_cap_tot = sum(amount_per_cap)) %>% # total amount per cap
    mutate(avg_med_income_house = mean(med_income_house, na.rm = TRUE)) %>% #average median income
    mutate(avg_poc_pct = mean(poc_pct)) %>%
    mutate(avg_rural_pct = mean(rural_pct)) %>%
    select(state_fips, fips, amount_per_cap_tot, avg_med_income_house, avg_poc_pct, avg_rural_pct) %>% # set up reduce down to one entry per county
    distinct() # reduce

invest_per_cap_dataset <- round(mean(myWorking_per_cap$amount_per_cap_tot),2)
```

The average investment per capita is `r invest_per_cap_dataset`. 


### chloroplast map of investment per capita in each county 

```{r}
# investigating format I need region and value in to use choropleth functions
# data(county.map)

# making dataset
myCounty_map_df <- myWorking_per_cap %>%
  mutate(region = as.numeric(fips)) %>%
  mutate(value = round(amount_per_cap_tot, 2)) %>%
  select(region, value)

# making plot
 myCounty_map <- county_choropleth(myCounty_map_df, num_colors = 5, legend = "Investment Per Capita")  + 
  labs(caption = str_wrap("\nFigure XX: There is a lot of within state variation in investment per capita on the East Coast, South, Midwest and the West Coast. The West has less varition, and counties benefit form a higher investment per capita. Maine and Vermont follow the same pattern as the West.", 120)) +
  theme(plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, size = 25),
        text=element_text(family="serif"))

ggsave(filename ="Figures/county_invest_per_cap.jpeg", myCounty_map, width = 11, height = 8)

myCounty_map



```



